Validity of a model for a cantilever two-dimensional beam

### 2024-04-07 files/Chapter5/2D_FEM.tex
plate model

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, the Finite Element Method (FEM) is applied to the models of this dissertation, presented in Chapter 1. The goal of this chapter is to obtain an algorithm that can be used to calculate the eigenvalues and eigenvectors for the models. This is in preparation of chapter 6, where the eigenvalues and eigenvectors of the models are compared.

In chapter 4, a method was discussed to obtain the eigenvalues and eigenvectors for the Timoshenko beam model. Therefore it is not necessary to apply the Finite Element Method on the Timoshenko model to obtain the eigenvalues and eigenvectors. The Finite Element Method is applied to the two-dimensional and three-dimensional models.

@@

Define the space
\begin{eqnarray*}
S^h & = & \textrm{span}\left(\left\{\delta_i \ | \ i = 1,2,...,k \right\} \right).
\end{eqnarray*}

@@

The matrices ${O}$ and ${O_f}$ are the zero matrices of the same size as $\mathbf{M}$ and ${M_f}$ respectively.

>>>>

\documentclass[../../main.tex]{subfiles}
\begin{document}
\section{Introduction}
In this chapter, the Finite Element Method (FEM) is applied to the models of this dissertation, presented in Chapter 1. The goal of this chapter is to obtain an algorithm that can be used to calculate the eigenvalues and eigenvectors for the models. This is in preparation of Chapter 6, where the eigenvalues and eigenvectors of the models are compared.

In Chapter 4, a method was discussed to obtain the eigenvalues and eigenvectors for the Timoshenko beam model. Therefore it is not necessary to apply the Finite Element Method on the Timoshenko model to obtain the eigenvalues and eigenvectors. The Finite Element Method is applied to the two-dimensional and three-dimensional models.

@@

Define the space
\begin{eqnarray*}
S^h & = & \textrm{span}\left\{\delta_i \ | \ i = 1,2,...,k \right\}.
\end{eqnarray*}

@@

The matrices ${O}$ and ${O_F}$ are the zero matrices of the same size as $\mathbf{M}$ and ${M_f}$ respectively.

### 2024-04-07 files/Chapter5/3D_FEM.tex
plate model

\subsubsection{Problem 3D-1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) - (Q,\phi) \label{3DB_20}
\end{align}
for all $\phi \in T(\Omega)$.\\

@@

\subsection{Galerkin approximation}
As mentioned in the previous section, section \ref{2d_FEM_G}, the body $\Omega$ needs to be descritised. For this is three-dimensional body, three-dimensional elements shapes are required. The elements used in this dissertation are rectangular prismatic elements. These elements are also known as `brick-shaped' elements and is described in \cite{Wu06}. This shape of element is a natural choice for a three-dimensional elastic body with a rectangular cross-section.

@@

For the this three-dimensional model, piecewise Hermite tri-cubic basis functions are used. Although this is more complex than using piecwise tri-linear basis functions, as mentioned in section \ref{2d_FEM_G}, the tri-cubic basis functions ensure faster convergence and also the derivatives of the solutions.

@@

Define the space
\begin{eqnarray*}
	S^h & = & \textrm{span}\left(\left\{\delta_i \ | \ i = 1,2,...,k \right\} \right)
\end{eqnarray*}

@@

\subsubsection{Problem 3D-1G}
Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
\begin{eqnarray*}
	(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I \cdot \phi_i)
\end{eqnarray*} for $i = 1,2,...,k$. $Q^I$ is scalar vector obtained after interpolating the function $Q$ over the rectangular grid $\Omega$. i.e. $Q^I_{i,j, h} = Q(x_i,x_j, x_h)$ for $i = 1,2,...,n_1$, $j = 1,2,...,n_2$ and $h = 1,2,...,n_3$.

@@

\subsubsection{FEM matrices}
\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{M}_{ij} & = & \int_{\Omega} \phi_j \phi_i \ dV \ \\
		{K_{11}}_{ij} & = & \int_{\Omega} \partial_1\phi_j \partial_1\phi_i \ dV \  \\
		{K_{12}}_{ij} & = & \int_{\Omega} \partial_1\phi_j \partial_2\phi_i \ dV \ \\
		{K_{13}}_{ij} & = & \int_{\Omega} \partial_1\phi_j \partial_3\phi_i \ dV \ 
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		{K_{22}}_{ij} & = & \int_{\Omega} \partial_2\phi_j \partial_2\phi_i \ dV \  \\
		{K_{23}}_{ij} & = & \int_{\Omega} \partial_2\phi_j \partial_3\phi_i \ dV \ \\
		{K_{33}}_{ij} & = & \int_{\Omega} \partial_3\phi_j \partial_3\phi_i \ dV \ \\
	\end{eqnarray*}
\end{minipage}
for $i,j = 1,2,...,k.$\\

@@

\subsubsection{Problem 3D-1ODE}
Find a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\ddot{\bar{u}} & = & K\bar{u} + M_{f}Q^I. \label{3D_M}
\end{eqnarray} With $\bar{u}$ in the form $\bar{u} = \partial^i_1u \partial^j_2u \partial^k_3u$ for $i,j,k = 0,1,2,3$.

>>>>

\subsubsection{Problem 3D-1V}
Find a function $u$ such that for all $t>0$, $u \in T(\Omega)$ and 
\begin{align}
	c(u,\phi) = -b(u,\phi) + (Q,\phi) \label{3DB_20}
\end{align}
for all $\phi \in T(\Omega)$.\\

@@

\subsection{Galerkin approximation}
As mentioned in the previous section, section \ref{2d_FEM_G}, the body $\Omega$ needs to be descritised. For this three-dimensional body, three-dimensional elements shapes are required. The elements used in this dissertation are rectangular prismatic elements. These elements are also known as `brick-shaped' elements and is described in \cite{Wu06}. This shape of element is a natural choice for a three-dimensional elastic body with a rectangular cross-section.

@@

For this three-dimensional model, piecewise Hermite tri-cubic basis functions are used. Although this is more complex than using piecwise tri-linear basis functions, as mentioned in section \ref{2d_FEM_G}, the tri-cubic basis functions ensure faster convergence and also the derivatives of the solutions.

@@

Define the space
\begin{eqnarray*}
	S^h & = & \textrm{span}\left\{\delta_i \ | \ i = 1,2,...,k \right\}
\end{eqnarray*}

@@

\subsubsection{Problem 3D-1G}
Find a function $u^h$ such that for all $t>0$, $u^h \in S^h$ and
\begin{eqnarray*}
	(u^h, \phi_i) & = & -b(u^h,\phi_i) + (Q^I \cdot \phi_i)
\end{eqnarray*} for $i = 1,2,...,k$. $Q^I$ is a scalar vector obtained after interpolating the function $Q$ over the rectangular grid $\Omega$, i.e. $Q^I_{i,j, h} = Q(x_i,x_j, x_h)$ for $i = 1,2,...,n_1$, $j = 1,2,...,n_2$ and $h = 1,2,...,n_3$.

@@

\subsubsection{FEM matrices}
\noindent\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		\mathbf{M}_{ij} & = & \int_{\Omega} \phi_j \phi_i \ dV \ \\
		{K_{11}}_{ij} & = & \int_{\Omega} \partial_1\phi_j \partial_1\phi_i \ dV \  \\
		{K_{12}}_{ij} & = & \int_{\Omega} \partial_1\phi_j \partial_2\phi_i \ dV \ \\
		{K_{13}}_{ij} & = & \int_{\Omega} \partial_1\phi_j \partial_3\phi_i \ dV \ 
	\end{eqnarray*}
\end{minipage}%
\begin{minipage}{.5\linewidth}
	\begin{eqnarray*}
		{K_{22}}_{ij} & = & \int_{\Omega} \partial_2\phi_j \partial_2\phi_i \ dV \  \\
		{K_{23}}_{ij} & = & \int_{\Omega} \partial_2\phi_j \partial_3\phi_i \ dV \ \\
		{K_{33}}_{ij} & = & \int_{\Omega} \partial_3\phi_j \partial_3\phi_i \ dV \ \\
	\end{eqnarray*}
\end{minipage}

for $i,j = 1,2,...,k.$\\

@@

\subsubsection{Problem 3D-1ODE}
Find a function $\bar{u} \in S^h$ such that
\begin{eqnarray}
	M\ddot{\bar{u}} & = & K\bar{u} + M_{f}Q^I. \label{3D_M}
\end{eqnarray} With $\bar{u}$ in the form $\bar{u} = \langle \partial^i_1u, \ \partial^j_2u, \ \partial^k_3u \rangle$ for $i,j,k = 0,1,2,3$.
